---
title: "make rules_allstudies"
author: "neele engelmann"
date: "2024-09-26"
output: html_document
---

# packages and data

```{r packages}
library(tidyverse)
library(here)
```

```{r get dataframes from study folders}
violate <- read.csv(here("violate","data","clean.csv"))
comply <- read.csv(here("comply","data","clean.csv"))
congruency_proportions <- read.csv(here("congruency_proportions","data","clean.csv"))
speed_accuracy <- read.csv(here("speed_accuracy","data","clean.csv")) 
stroop_flanker <- read.csv(here("stroop_flanker","data","clean.csv"))
punishment <- read.csv(here("punishment","data","clean.csv"))
```

# merge violate and comply

## some preparation: remove index column and practice trials, add column with study name

```{r prepare violate dataframe}
violate <- violate |> 
  select(-X) |> 
  mutate(study = rep("violate",length(violate$subj_code))) |> 
  filter(case!="noise") 
```

```{r prepare comply dataframe}
comply <- comply |> 
  select(-X) |> 
  mutate(study = rep("comply",length(comply$subj_code))) |> 
  filter(case!="noise")

#add congruent and condition variable to comply dataframe
comply <- comply |> 
  mutate(condition = case_when(
    text == 1 & purpose == 1 ~ "full violation",
    text == 1 & purpose == 0 ~ "text violation",
    text == 0 & purpose == 1 ~ "purpose violation",
    text == 0 & purpose == 0 ~ "no violation",
  ),
  congruent = ifelse(text==purpose,1,0))
```

## merge

```{r merge violate and comply}
rules_allstudies <- bind_rows(violate,comply)
```

<!-- # add the rules and behavior variables from codingsheet -->

<!-- ```{r import codingsheet} -->
<!-- codingsheet <- read.csv("codingsheet.csv",sep=";") -->

<!-- #only relevant variables -->
<!-- codingsheet <- codingsheet |> select(rule,question) -->
<!-- ``` -->

# ```{r add via fuzzyjoin with stimulus}
# library(fuzzyjoin)
# rules_allstudies <- regex_left_join(rules_allstudies, codingsheet, by = c("stimulus" = "question"))
# ```

note: this matches via behavor, not via rule. Slight deviations from rule phrasing in codingsheet and rule phrasing in stimulus variable are possible. But there is never a change in meaning. When in conflict, what participants actually saw is represented in "Stimulus".

- > try to achieve the same thing later with separate

# ```{r rename question to behavior}
# rules_allstudies <- rules_allstudies |> 
#   rename(behavior = question)
# ```

## order variables

```{r}
rules_allstudies <- rules_allstudies |> 
  select(study,subj_code,trial_index,case,stimulus,text,purpose,condition,congruent,response,rt)
```

# add congruency_proportions

## align varnames and levels for vars to keep

```{r add study variable}
congruency_proportions <- congruency_proportions |> 
  mutate(study = rep("congruency_proportions",length(congruency_proportions$subj_code))) 
```

```{r rename levels of scenario variable and rename variable}
congruency_proportions <- congruency_proportions |> 
  mutate(
    scen = case_when(
      scen == "auth" ~ "lab",
      scen == "cars" ~ "cars",
      scen == "dogs" ~ "restaurant",
      scen == "drin" ~ "alcohol",
      scen == "phon" ~ "phones",
      scen == "shoe" ~ "shoes",
      scen == "shoo" ~ "deer",
      scen == "slee" ~ "trainstation")) |> 
  rename(case = scen)
```

## remove vars not needed 

```{r remove type and conflict variable}
congruency_proportions <- congruency_proportions |> 
  select(-type,-conflict,-rule,-question)
```

## mark vars unique to this df

concerns block_order and conflict_prop. scen_number is also new but might be available for other dfs too

```{r}
congruency_proportions <- congruency_proportions |>
  rename(congprop_block_order = block_order,
        congprop = conflict_prop)
```

```{r more intelligible levels for block order variable}
congruency_proportions <- congruency_proportions |>
  mutate(congprop_block_order = ifelse(congprop_block_order==1,"high_low","low_high"))
```

## merge

```{r}
rules_allstudies <- bind_rows(rules_allstudies,congruency_proportions)
```

# add speed_accuracy

## align varnames and levels for vars to keep

```{r add study variable}
speed_accuracy <- speed_accuracy |> 
  mutate(study = rep("speed_accuracy",length(speed_accuracy$subj_code))) 
```

```{r rename stimtype to congruent}
speed_accuracy <- speed_accuracy |> 
  rename(congruent = stim_type)
```

```{r make congruent boolean}
speed_accuracy <- speed_accuracy |> 
  mutate(congruent = ifelse(congruent=="congruent",1,0))
```

```{r remove response and rename response_num to response}
speed_accuracy <- speed_accuracy |> 
  select(-response) |> 
  rename(response = response_num)
```

```{r rename correct answer variable}
speed_accuracy <- speed_accuracy |> 
  rename(correct_answer = correct_response)
```

```{r rename text and purpose violation}
speed_accuracy <- speed_accuracy |> 
  rename(text = text_violation,
         purpose = purpose_violation)
```

## remove vars not needed 

```{r remove yes key}
speed_accuracy <- speed_accuracy |> 
  select(-yes_key,-correct_answer,-wrong_answer)
```

## mark vars unique to this df

```{r rename speed accuracy condition order variable}
speed_accuracy <- speed_accuracy |> 
  rename(spd_acc_order = condition)
```

```{r rename mode variable}
speed_accuracy <- speed_accuracy |> 
  rename(spd_acc_mode = mode)
```

## merge

```{r}
rules_allstudies <- bind_rows(rules_allstudies, speed_accuracy)
```

# add stroop_flanker

## align varnames and levels for vars to keep

```{r create study variable}
stroop_flanker <- stroop_flanker |> 
  mutate(study = rep("stroop_flanker",length(stroop_flanker$subj_code))) 
```

```{r rename text and purpose vars}
stroop_flanker <- stroop_flanker |>
  rename(text = rules_textviolation,
         purpose = rules_purposeviolation)
```

```{r scen variable}
stroop_flanker <- stroop_flanker |> 
  mutate(
    rules_scen = case_when(
      rules_scen == "auth" ~ "lab",
      rules_scen == "cars" ~ "cars",
      rules_scen == "dogs" ~ "restaurant",
      rules_scen == "drin" ~ "alcohol",
      rules_scen == "phon" ~ "phones",
      rules_scen == "shoe" ~ "shoes",
      rules_scen == "shoo" ~ "deer",
      rules_scen == "slee" ~ "trainstation",
      rules_scen == NA ~ NA)) |> 
  rename(case = rules_scen)
```

```{r add condition variable}
stroop_flanker <- stroop_flanker |> 
  mutate(condition = case_when(
    text == 1 & purpose == 1 ~ "full violation",
    text == 1 & purpose == 0 ~ "text violation",
    text == 0 & purpose == 1 ~ "purpose violation",
    text == 0 & purpose == 0 ~ "no violation"))
```

```{r make congruent variable boolean}
stroop_flanker <- stroop_flanker |> 
  mutate(
    congruent = ifelse(congruent=="congruent",1,0)
  )
```

## mark vars unique to this df

dealing with response variables for the different tasks

* response should only indicate yes (1) and no (0) for rule violation judgments, so the variable can be easily compared between all studies

```{r}
stroop_flanker <- stroop_flanker |> 
  mutate(
    rules_response_verbal = case_when(
      rules_response_verbal == "yes" ~ 1,
      rules_response_verbal == "no" ~ 0,
      rules_response_verbal == NA ~ NA,
    )
  ) |> 
  rename(response = rules_response_verbal)
```

* response_stroop contains the response to the stroop task (y,b,r,g)

```{r}
stroop_flanker <- stroop_flanker |> 
  mutate(
    response_stroop = ifelse(task=="stroop",response_input,NA)
  )
```

* response_flanker contains the response to the flanker task (e,i)

```{r}
stroop_flanker <- stroop_flanker |> 
  mutate(
    response_flanker = ifelse(task=="flanker",response_input,NA)
  )
```

* stroopflanker_correct contains the correct response option for stroop and flanker task

```{r}
stroop_flanker <- stroop_flanker |> 
  mutate(
    stroopflanker_correct_resp = ifelse(stroopflanker_correct_resp=="",NA,stroopflanker_correct_resp)
  )

```

* correct_answer codes whether stroop and flanker have been answered correctly

```{r rename and make boolean}
stroop_flanker <- stroop_flanker |>
  mutate(stroopflanker_correct = case_when(
    stroopflanker_correct == "true" ~ 1,
    stroopflanker_correct == "false" ~ 0,
    stroopflanker_correct == "" ~ NA)) |> 
  rename(correct_answer = stroopflanker_correct)
```

## remove vars not needed

select only needed vars

```{r}
stroop_flanker <- stroop_flanker |> 
  select(
    study,
    subj_code,
    trial_index,
    task,
    stimulus,
    case,
    text,
    purpose,
    condition,
    congruent,
    response,
    response_stroop,
    response_flanker,
    rt,
    stroopflanker_correct_resp,
    correct_answer
  )
```


## merge

```{r}
rules_allstudies1 <- bind_rows(rules_allstudies,stroop_flanker)
```

# add punishment

## align varnames and levels

```{r delete row numbers}
punishment <- punishment |> 
  select(-X)
```

```{r add study name}
punishment <- punishment |> 
  mutate(study = rep("punishment",length(punishment$subj_code)))
```

```{r rename text and purpose vars}
punishment <- punishment |> 
  rename(text = text_violation,
         purpose = purpose_violation)
```

```{r create congruent}
punishment <- punishment |> 
  mutate(
    congruent = ifelse(text==purpose,1,0)
  )
```

```{r add condition variable}
punishment <- punishment |> 
  mutate(condition = case_when(
    text == 1 & purpose == 1 ~ "full violation",
    text == 1 & purpose == 0 ~ "text violation",
    text == 0 & purpose == 1 ~ "purpose violation",
    text == 0 & purpose == 0 ~ "no violation"))
```

## remove vars not needed

task
key_assignment
yeskey
scen
rule
question
conflict

```{r}
punishment <- punishment |> 
  select(-task,-key_assignment,-yeskey,-rule,-question,-conflict,-type)
```

## merge

```{r}
rules_allstudies <- bind_rows(rules_allstudies,punishment)
```

# final reorganization of dataframe

- create rules and behavior variables from codingsheet or from stimulus
- task variable should be rules for everything except stroop and flanker tasks in stroop_flanker
- correct answer: code textualist for rules task, correct/incorrect for stroop and flanker

# export

```{r}
write.csv(rules_allstudies,"rules_allstudies.csv",row.names=F)
```



