---
title: "stroop_flanker_analysis_behavioral"
format: html
editor: visual
---
```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE)
```

# packages and data

```{r}
library(here)
library(tidyverse)
```

```{r}
data_path <- here("stroop_flanker_raw.csv")
data <- read.csv(data_path)
```

# separate by task, clean

## stroop data

```{r}
stroop <- data |> 
  filter(task=="stroop_response") |> 
  select(subj_code,color,text,response,correct_response,correct,rt,trial_index) |> 
  mutate(congruent = ifelse(color==text,1,0))
```

correct responses for congruent vs. incongruent trials

```{r}
#calculate prop.s per condition
p_stroop_prop <- stroop |> 
  mutate(correct=factor(ifelse(correct=="true",1,0))) |> 
  group_by(congruent,correct) |> 
  summarize(n = n()) |> 
  mutate(prop=round(n/sum(n),2),
         se = sqrt((prop*(1-prop))/n),
         CI_lower = prop-1.96*se,
         CI_upper = prop+1.96*se) |> 
  filter(correct==1) |>
  mutate(congruent=ifelse(congruent==1,"congruent","incongruent")) |> 
  ggplot(aes(x=congruent,y=prop))+
    geom_bar(stat = "identity",aes(fill=congruent),width=0.5,alpha=0.6)+
    geom_linerange(aes(ymin=CI_lower,ymax=CI_upper))+
    theme_minimal()+
  labs(x=" ",y="prop. correct\n")+
  scale_fill_manual(values=c("#61A0AF","#F5D491"))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,1))+
  theme(axis.title.y = element_text(size=10,color="grey20"))

p_stroop_prop
```

response times for congruent vs. incongruent trials

```{r}
#medians
stroop_rt_medians <- stroop |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct=="true","correct responses","incorrect responses")),
         rt=as.numeric(rt)) |>
  group_by(congruent,correct) |> 
  summarize(m = median(rt))

stroop_rt_medians

#plot
p_stroop_rt <- stroop |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct=="true","correct responses","incorrect responses")),
         rt=as.numeric(rt)) |>
  ggplot(aes(x=rt,fill=factor(congruent)))+
  geom_density(alpha=0.6,color=NA)+
  facet_wrap(~correct)+
  geom_vline(data=stroop_rt_medians,aes(xintercept=m,color=factor(congruent)),linewidth=1)+
  theme_minimal()+
  scale_fill_manual(values=c("#F5D491","#61A0AF"))+
  scale_color_manual(values=c("#F5D491","#61A0AF"))+
  theme(legend.position = "none")+
  labs(x=" ",y="density\n")+
  theme(axis.title.y = element_text(size=10,color="grey20"))


p_stroop_rt
```
merge plots

```{r}
library(ggpubr)

p_stroop_merged <- ggarrange(p_stroop_prop,p_stroop_rt,widths = c(1, 3))
```

## flanker data

```{r}
flanker <- data |> 
  filter(direction=="left"|direction=="right") |> 
  select(subj_code,stimulus,direction,flanker_stim_type,correct_response,correct,rt,trial_index) |> 
  rename(congruent = flanker_stim_type) |> 
  mutate(correct=ifelse(correct=="true",1,0))

#check
flanker |> 
  group_by(stimulus,direction,congruent,correct_response) |>
  summarize(n=n())
```

correct responses for congruent vs. incongruent trials

```{r}
#calculate prop.s per condition
p_flanker_prop <- flanker |> 
  group_by(congruent,correct) |> 
  summarize(n = n()) |> 
  mutate(prop=round(n/sum(n),2),
         se = sqrt((prop*(1-prop))/n),
         CI_lower = prop-1.96*se,
         CI_upper = prop+1.96*se) |> 
  filter(correct==1) |>
  ggplot(aes(x=congruent,y=prop))+
    geom_bar(stat = "identity",aes(fill=congruent),width=0.5,alpha=0.6)+
    geom_linerange(aes(ymin=CI_lower,ymax=CI_upper))+
    theme_minimal()+
  labs(x=" ",y="prop. correct\n")+
  scale_fill_manual(values=c("#61A0AF","#F5D491"))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,1))+
  theme(axis.title.y = element_text(size=10,color="grey20"))


p_flanker_prop
```

response times for congruent vs. incongruent trials

```{r}
#medians
flanker_rt_medians <- flanker |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct==1,"correct responses","incorrect responses")),
         rt=as.numeric(rt)) |>
  group_by(congruent,correct) |> 
  summarize(m = median(rt))

flanker_rt_medians

#plot
p_flanker_rt <- flanker |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct==1,"correct responses","incorrect responses")),
         rt=as.numeric(rt)) |>
  ggplot(aes(x=rt,fill=factor(congruent)))+
  geom_density(alpha=0.6,color=NA)+
  facet_wrap(~correct)+
  geom_vline(data=flanker_rt_medians,aes(xintercept=m,color=factor(congruent)),linewidth=1)+
  theme_minimal()+
  scale_fill_manual(values=c("#F5D491","#61A0AF"))+
  scale_color_manual(values=c("#F5D491","#61A0AF"))+
  theme(legend.position = "none")+
  labs(x=" ",y="density\n")+
  theme(axis.title.y = element_text(size=10,color="grey20"))


p_flanker_rt
```

merge plots

```{r}
p_flanker_merged <- ggarrange(p_flanker_prop,p_flanker_rt,widths = c(1, 3))
```

## rules

```{r}
rules <- data |> 
  filter(task=="response") |> 
  select(subj_code,stimulus,key_assignment,response,rt) |> 
  mutate(case = ifelse(str_detect(stimulus,"dog"),"restaurant",
                       ifelse(str_detect(stimulus,"shoes"),"shoes",
                              ifelse(str_detect(stimulus,"dormitory"),"noise",
                                     ifelse(str_detect(stimulus,"train station"),"trainstation",
                                            ifelse(str_detect(stimulus,"parliament"),"deer",
                                                   ifelse(str_detect(stimulus,"zero tolerance"),"alcohol",
                                                          ifelse(str_detect(stimulus,"No cars are allowed"),"cars",
                                                                 ifelse(str_detect(stimulus,"headmaster"),"phones","lab")))))))
),
yeskey = ifelse(str_detect(key_assignment,"e for yes"),"e","i"),
response_num = ifelse(response==yeskey,1,0)) |> 
  select(-response) %>% 
  rename(response = response_num)

#condition coding and more reformatting
codingsheet <- read.csv("codingsheet.csv",sep=";") 

library(fuzzyjoin)

rules <- rules %>% 
  filter(case!="noise") %>% 
  regex_left_join(codingsheet, by = c(stimulus = "question")) |> 
  mutate(congruent=ifelse(conflict==1,0,1),
         correct = ifelse(response==text_violation,1,0))

#check
rules |> 
  group_by(case,type) |> 
  summarize(n = n())

```
correct responses for congruent vs. incongruent trials

```{r}
#calculate prop.s per condition
p_rules_prop <- rules |> 
  group_by(congruent,correct) |> 
  summarize(n = n()) |> 
  mutate(prop=round(n/sum(n),2),
         se = sqrt((prop*(1-prop))/n),
         CI_lower = prop-1.96*se,
         CI_upper = prop+1.96*se) |> 
  filter(correct==1) |>
  mutate(congruent=ifelse(congruent==1,"congruent","incongruent")) |> 
  ggplot(aes(x=congruent,y=prop))+
    geom_bar(stat = "identity",aes(fill=congruent),width=0.5,alpha=0.6)+
    geom_linerange(aes(ymin=CI_lower,ymax=CI_upper))+
    theme_minimal()+
  labs(x="\ntrial type",y="prop. textualist\n")+
  scale_fill_manual(values=c("#61A0AF","#F5D491"))+
  theme(legend.position = "none")+
  scale_y_continuous(limits=c(0,1))+
  theme(axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(size=10,color="grey20"))

p_rules_prop
```
response times for congruent vs. incongruent trials

```{r}
#medians
rules_rt_medians <- rules |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct==1,"textualist responses","nontextualist responses"),levels=c("textualist responses","nontextualist responses")),
         rt=as.numeric(rt)) |>
  group_by(congruent,correct) |> 
  summarize(m = median(rt))

rules_rt_medians

#plot
p_rules_rt <- rules |> 
  filter(rt!="null") |> #filter out timed-out trials
  mutate(correct=factor(ifelse(correct==1,"textualist responses","nontextualist responses"),levels=c("textualist responses","nontextualist responses")),
         rt=as.numeric(rt)) |>
  ggplot(aes(x=rt,fill=factor(congruent)))+
  geom_density(alpha=0.6,color=NA)+
  facet_wrap(~correct)+
  geom_vline(data=rules_rt_medians,aes(xintercept=m,color=factor(congruent)),linewidth=1)+
  theme_minimal()+
  scale_fill_manual(values=c("#F5D491","#61A0AF"))+
  scale_color_manual(values=c("#F5D491","#61A0AF"))+
  theme(legend.position = "none")+
  labs(x="\nreaction times",y="density\n")+
  theme(axis.title.x = element_text(face="bold"),
        axis.title.y = element_text(size=10,color="grey20"))

p_rules_rt
```

merge plots

```{r}
p_rules_merged <- ggarrange(p_rules_prop,p_rules_rt,widths = c(1, 3))
```

merge all plots

```{r}
p_all <- ggarrange(p_stroop_merged,p_flanker_merged,p_rules_merged,nrow=3,ncol=1,labels=c("Stroop\n","Flanker\n","Rules\n"),font.label = list(size = 10, color = "black", family = NULL),label.y=1.2)

p_all <- p_all+theme(plot.margin=margin(12,3,3,3,"mm"))

ggsave(p_all,file="plot_behavioral.pdf",width=23,height=15,units="cm")
```

