---
title: "analysis"
format: html
editor: visual
execute:
  warning: false
  message: false
---

# packages

```{r}
library(lmtest)
library(lme4)
library(nlme)
library(rcompanion)
library(reshape2)
library(here)
library(tidyverse)
```

# data

```{r}
clean <- read.csv(here("data","clean.csv"))
```

# plots

## theme

```{r}
mytheme <- theme_classic()+
           theme(axis.text.x = element_text(color="black",size=8),
                  axis.title.x = element_text(color="black",size=10,face="bold"),
                  axis.text.y = element_text(color="black",size=8),
                  axis.title.y = element_text(color="black",size=10,face="bold"),
                  strip.text.x = element_text(size = 8),
                  legend.title = element_text(size = 10),
                  title = element_text(color="black",size=10,face="bold"),
                  legend.position = "right")
```

## judgments

### plot

```{r}
p1 <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  ggplot(aes(x=response))+
  geom_bar(stat="count",aes(fill=response))+
  mytheme+
  facet_wrap(~purpose+text)+
  labs(subtitle="Did this person violate the rule?")+
  scale_fill_manual(values=c("#619CFF","#F8766D"))
  

p1
```

### save

```{r}
ggsave(p1,file=here("figures","judgments.png"),width=13,height=17,units="cm")
```

## rts

### plot

```{r}
p2 <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  ggplot(aes(x=rt))+
  geom_density(aes(fill=response,color=response),alpha=0.5)+
  facet_wrap(~text+purpose)+
  mytheme+
  labs(subtitle="Did this person violate the rule?")+
  scale_fill_manual(values=c("#619CFF","#F8766D"))+
  scale_color_manual(values=c("#619CFF","#F8766D"))

p2

#median rts for overlaying
median_rts <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  group_by(text,purpose,response) %>% 
  summarize(rt = median(rt))

#overlay medians

p2 <- p2 + geom_vline(data=median_rts,aes(xintercept=rt,color=response),linewidth=1)

p2
```

### save

```{r}
ggsave(p2,file=here("figures","rts.png"),width=13,height=17,units="cm")
```

# preregistered analyses

## judgments

### descriptives

```{r}
clean |> 
  group_by(condition,response) |> 
  summarize(n = n()) |> 
  mutate(prop = round(n/sum(n),2),
         se = sqrt(prop*(1-prop)/n),
         CI_lower = round(prop - 1.96*se,2),
         CI_upper = round(prop + 1.96*se,2))
```

### logistic regression

"1) A mixed-effects logistic regression of rule violation judgments, with text, purpose and the text\*purpose interactions as fixed effects (and scenarios and participants as crossed, random effects). - We predict main (i.e., positive) effects of text and purpose."

```{r}
# null model
null <- glmer(response ~ 1 + (1|subj_code) + (1|case), data = clean,na.action=na.exclude, family=binomial)

# main effect text

purpose_only <- glmer(response ~ purpose + (1|subj_code) + (1|case), data = clean,na.action=na.exclude, family=binomial)

purpose_text <- glmer(response ~ purpose + text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(purpose_only,purpose_text) # better with text

# main effect purpose

text_only <- glmer(response ~ text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(text_only,purpose_text) # better with purpose

# interaction

interaction <- glmer(response ~ purpose * text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(purpose_text, interaction) # not better

# final model

finalmodel_judgments <- purpose_text

summary(finalmodel_judgments)

# odds ratios from coefficients

# effect of purpose violation:

exp(2.23594) # -> more likely that people say yes to violate question

# effect of text violation:

exp(3.96893) # -> more likely that people say yes to violate question, stronger effect than for purpose

nagelkerke(finalmodel_judgments,null)
```

## rts

"2) A mixed-effects ordinal regression of response times, with text, purpose and the textx purpose interactions as fixed effects (and scenario and participant as crossed random effects). -We predict an interaction effect of text\*purpose (indicating longer response times on conflict trials)."

### descriptives

```{r}
clean |> 
  group_by(condition) |> 
  summarize(median = median(rt))
```

```{r}
clean |> 
  group_by(response) |> 
  summarize(median = median(rt))
```

```{r}
clean |> 
  group_by(condition,response) |> 
  summarize(median = median(rt))
```

### linear model

```{r}
# null model
null <- glmer(rt ~ 1 + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

# main effect text

purpose_only <- lmer(rt ~ purpose + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

purpose_text <- lmer(rt ~ purpose + text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(purpose_only,purpose_text) # better with text

# main effect purpose

text_only <- lmer(rt ~ text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(text_only,purpose_text) # better with purpose

# interaction text x purpose

interaction <- lmer(rt ~ purpose*text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(purpose_text,interaction) # better
```

### interaction with response

```{r}
interaction_resp <- lmer(rt ~ purpose*text*response + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(interaction, interaction_resp)
```
