---
title: "analysis violate full sample"
output: 
  html_document:
    toc: true 
    toc_depth: 3  
    number_sections: true  
    theme: united 
    highlight: tango  
date: "2023-01-26"
---

# settings

```{r echo = F, include = F}
knitr::opts_chunk$set(message=F,warning=F)
```

# packages and data import

```{r import and cleaning}
library(tidyverse)
library(readr)
library(reshape2)

#raw data
raw_batch1 <- read.csv("raw_batch1.csv")
raw_batch2 <- read.csv("raw_batch2.csv")

#make unique IDs for batch 1
library(stringi)

ids_raw_batch1 <- raw_batch1 %>% 
  group_by(run_id) %>% 
  summarize(max = max(run_id)) %>% 
  rowwise() %>% 
  mutate(id = paste0(stri_rand_strings(1, 5, "[A-Z]"), stri_rand_strings(1, 4, "[0-9]"), stri_rand_strings(1, 1, "[A-Z]"))) %>% 
  select(-max)

raw_batch1 <- left_join(raw_batch1,ids_raw_batch1,by="run_id")

nlevels(factor(raw_batch1$id))

#make unique IDs for batch 2
ids_raw_batch2 <- raw_batch2 %>% 
  group_by(run_id) %>% 
  summarize(max = max(run_id)) %>% 
  rowwise() %>% 
  mutate(id = paste0(stri_rand_strings(1, 5, "[A-Z]"), stri_rand_strings(1, 4, "[0-9]"), stri_rand_strings(1, 1, "[A-Z]"))) %>% 
  select(-max)

raw_batch2 <- left_join(raw_batch2,ids_raw_batch2,by="run_id")

nlevels(factor(raw_batch2$id))

#merge
raw <- rbind(raw_batch1,raw_batch2)

nlevels(factor(raw$id))

raw <- raw %>% 
  mutate(subj_code = id)

nlevels(factor(raw$subj_code))
```

# cleaning and reformatting

```{r}
clean <- raw %>% 
  select(subj_code,stimulus,task,rt,key_assignment,trial_index,response) %>% 
  filter(task=="response") %>% 
  mutate(case = ifelse(str_detect(stimulus,"dog"),"restaurant",
                       ifelse(str_detect(stimulus,"shoes"),"shoes",
                              ifelse(str_detect(stimulus,"dormitory"),"noise",
                                     ifelse(str_detect(stimulus,"train station"),"trainstation",
                                            ifelse(str_detect(stimulus,"parliament"),"deer",
                                                   ifelse(str_detect(stimulus,"zero tolerance"),"alcohol",
                                                          ifelse(str_detect(stimulus,"No cars are allowed"),"cars",
                                                                 ifelse(str_detect(stimulus,"headmaster"),"phones","lab")))))))),
         yeskey = ifelse(str_detect(key_assignment,"e for yes"),"e","i"),
         response_num = ifelse(response==yeskey,1,0)) %>% 
  select(-response) %>% 
  rename(response = response_num)

codingsheet <- read.csv("codingsheet.csv") 

clean <- clean %>% 
  left_join(codingsheet,by="stimulus") %>% 
  select(subj_code,case,stimulus,text,purpose,trial_index,response,rt)

#check number of participants
nlevels(factor(clean$subj_code))

#check trials per participant
clean %>% 
  dplyr::group_by(subj_code) %>% 
  dplyr::summarize(n = n())

#exclude noncompleters
noncompleters <- clean %>% 
  dplyr::group_by(subj_code) %>% 
  dplyr::summarize(ntrials = n()) %>% 
  filter(ntrials < 108)

clean <- clean %>% dplyr::filter(!subj_code %in% noncompleters$subj_code)

#exclude rts under 200 ms, over 800 ms, and invalid trials (no response given)

clean <- clean %>% 
  filter(rt!="null") %>% # exclude invalid responses
  mutate(rt = as.numeric(rt)) %>% # convert to numeric
  dplyr::filter(rt > 200 & rt < 8000) # exclude very fast and very slow responses

#check distribution of rts after exclusions

summary(clean$rt)

#check new number of participants
nlevels(factor(clean$subj_code))

#check new number of trials per participant
clean %>% 
  dplyr::group_by(subj_code) %>% 
  dplyr::summarize(ntrials = n()) %>% 
  summary(ntrials)

#condition variable for hddm
clean <- clean %>% 
  mutate(condition = case_when(
    text==0 & purpose==0 ~ "no violation",
    text==1 & purpose==0 ~ "text violation",
    text==0 & purpose==1 ~ "purpose violation",
    text==1 & purpose==1 ~ "full violation"
  ),
  congruent = ifelse(text==purpose,1,0))

#save final file
write.csv(clean,"clean.csv")
```

# judgments by condition

## plot

```{r}
mytheme <- theme_classic()+
           theme(axis.text.x = element_text(color="black",size=8),
                  axis.title.x = element_text(color="black",size=10,face="bold"),
                  axis.text.y = element_text(color="black",size=8),
                  axis.title.y = element_text(color="black",size=10,face="bold"),
                  strip.text.x = element_text(size = 8),
                  legend.title = element_text(size = 10),
                  title = element_text(color="black",size=10,face="bold"),
                  legend.position = "right")
```

```{r}
p1 <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  ggplot(aes(x=response))+
  geom_bar(stat="count",aes(fill=response))+
  mytheme+
  facet_wrap(~purpose+text)+
  labs(subtitle="Did [agent] violate the rule?")+
  scale_fill_manual(values=c("#619CFF","#F8766D"))
  

p1

ggsave(p1,file="plot_judgments.png",width=13,height=17,units="cm")
```

## logistic regression

```{r}
library(lmtest)
library(lme4)
library(nlme)
library(rcompanion)
```

```{r}
# null model
null <- glm(response ~ 1, data = clean,na.action=na.exclude, family=binomial)

# random intercept for participant
RI_id <- glmer(response ~ 1 + (1|subj_code), data = clean,na.action=na.exclude, family=binomial)

lrtest(null,RI_id) #not better

# random intercept for scenario
RI_scenario <- glmer(response ~ 1 + (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(null,RI_scenario) # better 

# main effect text

purpose_only <- glmer(response ~ purpose+ (1|case), data = clean,na.action=na.exclude, family=binomial)

purpose_text <- glmer(response ~ purpose + text+ (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(purpose_only,purpose_text) # better with text

# main effect purpose

text_only <- glmer(response ~ text + (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(text_only,purpose_text) # better with purpose

# interaction

interaction <- glmer(response ~ purpose * text+ (1|case), data = clean,na.action=na.exclude, family=binomial)

lrtest(purpose_text, interaction) # not better

# final model

finalmodel_judgments <- purpose_text

summary(finalmodel_judgments)

# odds ratios from coefficients

# effect of purpose violation:

exp(2.20354) # -> more likely that people say yes to violate question

# effect of text violation:

exp(3.88776) # -> more likely that people say yes to violate question, stronger effect than for purpose

nagelkerke(finalmodel_judgments,null)

```

## plot learning effects

```{r}
p1_learning <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    )) %>% 
  ggplot(aes(x=trial_index,y=response)) + 
  geom_smooth(method="lm",aes(group=subj_code),se=F,linewidth=0.5,color="grey65")+
  geom_smooth(method="lm",se=F,linewidth=0.8,color="black")+
  facet_wrap(~text+purpose)+
  mytheme

p1_learning

ggsave(p1_learning,file="plot_judgments_learning.png",width=13,height=17,units="cm")
```

## analysis learning effects

linear regression per condition - significant slope for trial index?

```{r}
summary(lm(response~trial_index,data=clean %>% filter(text==0 & purpose == 0))) #no

summary(lm(response~trial_index,data=clean %>% filter(text==1 & purpose == 0))) #yes, positive (people become more likely to say yes to violate question)

summary(lm(response~trial_index,data=clean %>% filter(text==0 & purpose == 1))) #yes, negative (people become less likely to say yes to violate question)

summary(lm(response~trial_index,data=clean %>% filter(text==1 & purpose == 1))) #yes, positive (people become more likely to say yes to violate question)
```

# reaction times by condition and response (yes vs. no)

## plot

```{r}
p2 <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  ggplot(aes(x=rt))+
  geom_density(aes(fill=response,color=response),alpha=0.5)+
  facet_wrap(~text+purpose)+
  mytheme+
  labs(subtitle="Did [agent] violate the rule?")+
  scale_fill_manual(values=c("#619CFF","#F8766D"))+
  scale_color_manual(values=c("#619CFF","#F8766D"))

p2

#median rts for overlaying
median_rts <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"
    ),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")
    ) %>% 
  group_by(text,purpose,response) %>% 
  summarize(rt = median(rt))

#overlay medians

p2 <- p2 + geom_vline(data=median_rts,aes(xintercept=rt,color=response),linewidth=1)

p2

ggsave(p2,file="plot_rts.png",width=13,height=17,units="cm")

```

## linear model

predictions:

- main effect response (yes faster than no)

- interaction text x purpose (longer rts for conflict trials)

```{r}
#null
null <- glm(rt ~ 1, data = clean,na.action=na.exclude)

#ri id
ri_id <-lmer(rt ~ 1 + (1|subj_code), data = clean,na.action=na.exclude)

lrtest(null,ri_id) #better

#ri scenario

ri_id_scen <- lmer(rt ~ 1 + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(ri_id,ri_id_scen) #better

#main effect text
purpose_response <- lmer(rt ~ purpose + response + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

full <- lmer(rt ~ purpose + response + text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(purpose_response,full) #better

#main effect purpose

text_response <- lmer(rt ~ text + response + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(text_response,full) #better

#main effect response

text_purpose <- lmer(rt ~ text + purpose + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(text_purpose,full) #better

#interaction text x purpose

interaction <- lmer(rt ~ purpose * response + text + (1|subj_code) + (1|case), data = clean,na.action=na.exclude)

lrtest(full,interaction) #better

#final model

finalmodel_rts <- interaction

summary(finalmodel_rts)

nagelkerke(finalmodel_rts,null)
```

## plot learning effects

```{r}
p2_learning <- clean %>% 
  mutate(text=case_when(
    text==0 ~ "no text violation",
    text==1 ~ "text violation"
      ),
    purpose=case_when(
      purpose==0 ~ "no purpose violation",
      purpose==1 ~ "purpose violation"),
    response=case_when(
      response==0 ~ "no",
      response==1 ~ "yes")) %>% 
  ggplot(aes(x=trial_index,y=rt)) + 
  geom_smooth(method="lm",se=T,linewidth=0.8,aes(group=response,color=response))+
  facet_wrap(~text+purpose)+
  mytheme+
  labs(subtitle="Did [agent] violate the rule?")+
  scale_fill_manual(values=c("#619CFF","#F8766D"))+
  scale_color_manual(values=c("#619CFF","#F8766D"))

p2_learning

ggsave(p2_learning,file="plot_rts_learning.png",width=13,height=17,units="cm")
```